---
title: "French given names per year per department"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "October, 2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Build the Dataframe from file

```{r}
library(tidyverse)
first_names <- read_delim("dpt2021.csv",delim=";", show_col_types = FALSE)
```

To make this analysis easier, I'll remove useless values such as "_PRENOMS_RARES" from the dataset. In addition, The year field will be converted to a proper date format, which should make visualization clearer with GGPLOT.

## Drop useless values and fix formatting in the dataframe
```{r}
first_names = subset(first_names, preusuel != '_PRENOMS_RARES' & annais != 'XXXX')
first_names <- first_names %>% rename(name = preusuel, year = annais, quantity = nombre)
first_names$year <- as.POSIXct(first_names$year, format="%Y")
```

# Analysis

## 1. Choose a firstname and analyse its frequency along time
I'm considering this just a warm-up for the next sections. Let's analyse the popularity of the name Mario over time:

```{r}
mario_popularity = first_names %>%
                  filter(name == "MARIO") %>%
                  group_by(name, year) %>%
                  summarise_at(vars(quantity), list(quantity = sum))

ggplot(data=mario_popularity, aes(x=year, y=quantity, group=1)) +
  geom_line() +
  geom_point()
```
Even at its most popular, the first name Mario didn't really have an impressive frequency. In section 2, we will see that the most popular names have much higher frequencies.

## 2. Most popular first names by gender
## Male
```{r}
popular_male_names = first_names %>%
                  filter(sexe == 1) %>% 
                  group_by(year, name) %>% 
                  summarise(popularity = sum(quantity)) %>% 
                  filter(popularity == max(popularity))
```

Now let's see the progression of those names over time, starting from 1960. What we want to analyze here is whether those recent popular names had already been popular sometime in the past and also whether their popularity is short-lived or not.
```{r}
most_popular_last_30y = tail(popular_male_names$name, 30)

top30_over_time = first_names %>%
                  filter(name %in% most_popular_last_30y & year > as.POSIXct("1960-01-01")) %>% 
                  group_by(year, name) %>% 
                  summarise(quantity_total = sum(quantity))

ggplot(data=top30_over_time, aes(x=year, y=quantity_total, color = name, group=name)) +
  geom_line() +
  geom_point()
```

It seems to be consistent that the most popular names tend to come in waves. Not only that, we cab also note that, over the last 30 years, only 8 male names have taken the top spot. Let's verify if the same phenomenon occurs for female names:

## Female
```{r}
popular_female_names = first_names %>%
                  filter(sexe == 2) %>% 
                  group_by(year, name) %>% 
                  summarise(popularity = sum(quantity)) %>% 
                  filter(popularity == max(popularity))
```

```{r}
most_popular_last_30y = tail(popular_female_names$name, 30)

top30_over_time = first_names %>%
                  filter(name %in% most_popular_last_30y & year > as.POSIXct("1960-01-01")) %>% 
                  group_by(year, name) %>% 
                  summarise(quantity_total = sum(quantity))

ggplot(data=top30_over_time, aes(x=year, y=quantity_total, color = name, group=name)) +
  geom_line() +
  geom_point()
```

Here we see the same pattern; even fewer first names have taken the top spot for female names. It can be noted, however, that the female names that have dominated over the last 30 years seem to have risen to popularity more recently than their male counterparts.

# 3. Optional : Which department has a larger variety of names along time ? Is there some sort of geographical correlation with the data?