---
output:
  pdf_document: default
  html_document: default
---
```{r}
library(ggplot2)
```
# Loading data
```{r}
myData=read.table(file="arbres-tot.csv",sep=";",skip=3,header=TRUE)
myData=myData[myData$X10!=0,]
```
# 1. Simple regression
```{r}
circ = myData$X70
height = myData$X10
simple_reg <- lm(height~circ,data=myData)
```

Plot the data to get an initial view of the possible correlation.

```{r}
ggplot(myData,aes(x=circ,y=height))+ geom_point()+
xlab("circ")+
ylab("height")
```

```{r}
anova(simple_reg)
summary(simple_reg)
```
Plot the regression line
```{r}
ggplot(myData,aes(x=circ,y=height))+ geom_point()+
stat_smooth(method="lm",se=FALSE)+ xlab("circ")+
ylab("height")
```


```{r}
plot(simple_reg$residuals)
plot(simple_reg,3)
plot(simple_reg,1)
plot(simple_reg,4)
```

```{r}
predict(simple_reg, data.frame(circ=10),interval="prediction")
```

# 2. Multivariate regression
Add a column to the sample. Name it circ_sqrt and fill it with the square root of the circumference of each
tree
```{r}
myData$circ_sqrt <- sqrt(myData$X70)
```

Perform the multivariate linear regression of height on the basis of:
• of circumference ;
• of circ_sqrt.



```{r}
multi_reg <- lm(height~circ+circ_sqrt,data=myData)
summary(multi_reg)
```
The variable **circ** is not significant at the 5% test level, so we will remove it for the regression.


```{r}
multi_reg_2 <- lm(height~circ_sqrt,data=myData)
summary(multi_reg_2)
```

Plot the regression

```{r}
circ_pred <- seq(0,175,len=1000)
height_pred <- multi_reg_2$coefficients[1]+multi_reg_2$coefficients[2]*sqrt(circ_pred)
fct_reg <- data.frame(circ_pred=circ_pred,height_pred=height_pred)
ggplot() + geom_point(data=myData,aes(x=circ,y=height)) + geom_line(data=fct_reg,aes(x=circ_pred,y=height_pred),col="blue") + stat_smooth(method="lm",se=FALSE) + xlab("circ") + ylab("height")
```