---
output:
  pdf_document: default
  html_document: default
---
# Library loading
```{r}
library(tidyverse)
library(forecast)
library(tseries)
```


# Loading data from the Mauna Loa observatory
```{r}
co2_monthly_full <- read.delim(file = 'co2_mm_mlo.txt', comment.char = '#', header = F, sep = '')
co2_monthly_full <- co2_monthly_full[, c(1, 2, 4)]
names(co2_monthly_full) <- c('year', 'month', 'co2_concentration')
head(co2_monthly_full)
```

# Data cleanup and transformations
Let's first check if the dataset contains null values
```{r}
which(is.na(co2_monthly_full))
```
Perfect, we don't have any null data. This is because the months in the downloaded dataset had already been interpolated.

Readings for years 1958 and 2022 will be removed since they do not have the 12 months included.
```{r}
co2_monthly_full <- co2_monthly_full %>% filter(year != 2022 & year != 1958)
```

We will also create a single date field based on month and year, which should improve the readability of our plots in the future.
```{r}
co2_monthly_full$date <- as.Date(paste(co2_monthly_full$year, co2_monthly_full$month, 1, sep = '-'), format = '%Y-%m-%d')
co2_monthly <- co2_monthly_full[, c('date', 'co2_concentration')]
head(co2_monthly)
```




Finally, we'll transform the dataframe into a time series object, which enables us to use some helpful R functions further in the analysis.
```{r}
co2_ts <- ts(co2_monthly$co2_concentration, start = c(1959, 1), frequency = 12)
```

# 1. Exploratory analysis
Let's plot the data for the first time:
```{r}
plot(
  co2_monthly$date,
  co2_monthly$co2_concentration,
  type = 'l',
  xlab = 'Date',
  ylab = 'CO2 Concentration PPM',
  main = 'Mauna Loa Weekly Carbon Dioxide Concentration'
)
```

From a purely visual analysis, it seems like the data presents both seasonality and an upward trend. Let's first take a look at the seasonality:

## Examining seasonality
```{r}
boxplot(co2_ts~cycle(co2_ts),xlab="Date", ylab = "CO2 concentration (PPM)",main ="Monthly CO2 average from 1959 to 2021")
```
Here we can see that the monthly concentrations show seasonality, with a high point around the months 4-5 and a low point around the months 9-10.

Let's take a look at the seasonality over the years:
```{r}
ggplot(data = co2_monthly_full, aes(factor(month), co2_concentration, colour = year, group = year)) +
  geom_line() +
  xlab('Month') +
  ylab('CO2 Concentration PPM') + 
  ggtitle('Mauna Loa Monthly Carbon Dioxide Concentration') +
  scale_color_gradientn('Year', colors = rainbow(length(unique(co2_monthly_full$month))))
```
The results seem to be consistent over the years, with the seasonality being maintained through the upward trend in CO2 concentrations.

From these plots we could gather that:
- The CO2 concentration shows a clear upward linear trend over the years
- The monthly concentrations show seasonality, with a peak around the months 4-5 and a valley around the months 9-10.

## Series decomposition
Let's take a better look at the time series decomposition:
```{r}
decomposeCO2 <- decompose(co2_ts,"additive")
autoplot(decomposeCO2)
```

Removing the trend seen above can be useful for further analysis. Let's check if differencing would be enough to make the time series stationary with adf.test():
```{r}
adf.test(diff(log(co2_ts)), alternative="stationary", k=0)
```
This low P value confirms the alternative hypothesis (that is, the time series has become stationary after one stage of differencing); now let's go ahead and plot ACF:
```{r}
diff(log(co2_ts)) %>% ggtsdisplay()
```

# Making predictions
## First attempt: 
Let's first take a look at the P value for the trend, using the date as the independent variable.
```{r}
simple_reg_co2 <- lm(co2_concentration ~ date, data = co2_monthly)
```

```{r}
ggplot(data = co2_monthly, aes(x = date, y = co2_concentration)) +
  geom_line() +
  xlab('Date') +
  ylab('CO2 Concentration PPM') + 
  ggtitle('Mauna Loa Weekly Carbon Dioxide Concentration') +
  stat_smooth(method = "lm", se = FALSE)
```
Visually, we can see that this model might not be perfect as it cannot follow the seasonality. I suspect the errors will have a high autocorrelation because of that:

```{r}
checkresiduals(simple_reg_co2)
```
As expected, the autocorrelation of errors is pretty high, which is not what we want from a forecasting model. Let's try another one.

## Second attempt: ARIMA
```{r}
arima_co2 <- auto.arima(co2_ts)
```
```{r}
checkresiduals(arima_co2)
```
Much better. auto.arima() is able to detect seasonality and adapt to it through statistical tests; now we can see that the errors follow a noise (random) pattern, which is preferable for a model.

### Sample forecast:
Now we can use the forecast function to plot a forecast on the horizon of 100 months, with a confidence interval of 95%.
```{r}
arima_forecast_co2 <- forecast(arima_co2, level = c(95), h = 100)
autoplot(arima_forecast_co2)
```

